---
title: "NYC Robberies Occur Near Schools at 3PM"
author: "Matt Moocarme"
date: "July 8, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      cache = T,
                      message = FALSE,
                      comment = '')
```

## R Markdown

Working from Ben Wellington's IQuantNY

Load in libraries
```{r }
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggmap)
library(caret)
```

Next we read in the data
```{r}
felony.data <- read_csv('data/NYPD_7_Major_Felony_Incidents.csv')
str(felony.data)
```
Tidy the data to look at weekday robberies via hour 
```{r}
tidy.fdata <- felony.data %>% filter(Offense == 'ROBBERY') %>% 
  select(Hour = `Occurrence Hour`,Day.Week = `Day of Week`) %>% 
  filter(!(Day.Week %in% c('Saturday', 'Sunday'))) %>%
  group_by(Hour) %>% summarise(Hourly.Crime.Rate= n())

```

Recreate plot from IQuantNY
```{r}
# Recreate plot from IQuantNY (roughly)
ggplot(data = tidy.fdata) + geom_line(aes(x = Hour, y = Hourly.Crime.Rate)
                                      , color= 'darkgreen')
```

Analyze data just at 3pm
```{r}
new.fdata <- felony.data %>% filter(Offense == 'ROBBERY') %>% 
  select(Hour = `Occurrence Hour`,Day.Week = `Day of Week`, `Location 1`) %>% 
  filter(!(Day.Week %in% c('Saturday', 'Sunday')), Hour =='15')

```
Read in school data
```{r}
# Read in school data ================================================
schooldf <- read_csv('data/Location_Information_Report.csv')
str(schooldf)
unique(schooldf$`Location Category Description`)
```
Since I don't believe elementary schools, pre-K re commiting robberies I will just use those that schools that contain the elder students
```{r}
rel.schools <- c('High School', 'Secondary School', 'K-12 all grades')

tidy.schooldf <- schooldf %>% select(`Location Name`, `Location Category Description`, Latitude, Longitude) %>%
  filter(`Location Category Description` %in% rel.schools)
dim(tidy.schooldf)
```

Get long/lat coords from the felony data use regualr expressions
```{r}
new.fdata2 <- new.fdata %>%
  separate(`Location 1`, c('Lat_tmp', 'Long_tmp'),sep = ',') %>%
  mutate(Lat = as.numeric(gsub("\\(|\\)","", Lat_tmp))) %>%
  mutate(Long = as.numeric(gsub("\\(|\\)","", Long_tmp)))

```
Cluster the coords- choose 150 clusters - more than the number of schools
```{r}
set.seed(666)
latlong_cluster <- kmeans(cbind(new.fdata2$Lat, new.fdata2$Long), centers  = 150)
```

bind new lat long coords, group by and summarise
```{r}
new.fdata3 <- cbind(new.fdata2, Cluster = latlong_cluster$cluster, 
                    Cluster.Lat = latlong_cluster$centers[latlong_cluster$cluster,1],
                    Cluster.Long = latlong_cluster$centers[latlong_cluster$cluster,2])

new.fdata4 <- new.fdata3 %>% group_by(Cluster.Lat, Cluster.Long, Cluster) %>% summarise(Total.Counts= n())

```
Plot on map

```{r}
# = Plot on Map
map <- get_map(location = 'New York City', zoom= 10,
               source = 'google', color = 'color')
map <- ggmap(map) + geom_point(data = new.fdata4, aes(x=Cluster.Long, y=Cluster.Lat, 
                                                      color = new.fdata4$Total.Counts), 
                               size = new.fdata4$Total.Counts/100) +
  scale_color_gradient(low = 'red', high = 'yellow') + 
  geom_jitter(data = tidy.schooldf, aes(x= Longitude, y= Latitude), shape = 4) +
  labs(title = 'New York City', x = 'Longitude', y = 'Latitude') 
map
```

